Sleep, relative to wake, increases both veridical and false memory in the DRM paradigm: A registered report

### 2023-07-25 manuscript.Rmd
Reverting to using pre-processed data as reached the 6 hour limit
Test autogenerate of the procedure figure

processdata <- 1   # running all analyses from scratch takes about 6 hours
buildfigures <- 1

packagelist <- c('bookdown','knitr','rmarkdown','osfr','plyr','tidyr','dplyr','emmeans','ggplot2','Rmisc','buildmer','lme4','lmerTest','stringr','effects','effectsize','ggdist','car','gridExtra','kableExtra','rstatix','simr','effsize','readr','LSAfun')
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

``` {r getdata}

if (!dir.exists('data')){dir.create('data')}
if (!dir.exists('Figures')){dir.create('Figures')}

# download the data if not already present
osfproject <- osf_retrieve_node("9pdyf")
osffiles <- osf_ls_files(osfproject,path='Stage 2/Data/All the csv needed for Final_Analysis.Rmd')
if (!file.exists('data/full.csv')){
  hid <- pmatch('full.csv',as.character(unlist(osffiles[,1])))
  osf_download(osffiles[hid,], path='data/')
}
if (!file.exists('data/lure_final.csv')){
  hid <- pmatch('lure_final.csv',as.character(unlist(osffiles[,1])))
  osf_download(osffiles[hid,], path='data/')
}
if (!file.exists('data/studied_final.csv')){
  hid <- pmatch('studied_final.csv',as.character(unlist(osffiles[,1])))
  osf_download(osffiles[hid,], path='data/')
}
if (!file.exists('data/Table 2_Material.csv')){
  hid <- pmatch('Table 2_Material.csv',as.character(unlist(osffiles[,1])))
  osf_download(osffiles[hid,], path='data/')
}

@@

>>>>

processdata <- 0   # running all analyses from scratch takes about 6 hours
buildfigures <- 1

packagelist <- c('bookdown','knitr','markdown','osfr','plyr','tidyr','dplyr','emmeans','ggplot2','Rmisc','buildmer','lme4','lmerTest','stringr','effects','effectsize','ggdist','car','gridExtra','kableExtra','rstatix','LSAfun','simr','effsize','readr', 'Gmisc', 'glue', 'htmlTable', 'grid', 'magrittr', 'tidyverse', 'magick', 'cowplot')
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

if (!dir.exists('data')){dir.create('data')}
if (!dir.exists('Figures')){dir.create('Figures')}
```

```{r getdata, include=FALSE}
# download the data if not already present
osfproject <- osf_retrieve_node("9pdyf")
osffiles <- osf_ls_files(osfproject,path='Stage 2/Data/All the csv needed for Final_Analysis.Rmd')
if (!file.exists('data/full.csv')){
  hid <- pmatch('full.csv',as.character(unlist(osffiles[,1])))
  osf_download(osffiles[hid,], path='data/')
}
if (!file.exists('data/lure_final.csv')){
  hid <- pmatch('lure_final.csv',as.character(unlist(osffiles[,1])))
  osf_download(osffiles[hid,], path='data/')
}
if (!file.exists('data/studied_final.csv')){
  hid <- pmatch('studied_final.csv',as.character(unlist(osffiles[,1])))
  osf_download(osffiles[hid,], path='data/')
}
if (!file.exists('data/Table 2_Material.csv')){
  hid <- pmatch('Table 2_Material.csv',as.character(unlist(osffiles[,1])))
  osf_download(osffiles[hid,], path='data/')
}

@@

```{r expprocedure-pdf, include=FALSE}

if (buildfigures==1){
#make boxes
immediate_group <- boxGrob(glue("Immediate Groups"),
                           txt_gp = gpar(col = "black", fontsize = 16),
                           box_gp = gpar(fill = 0, col = 0),
                           x = .25,
                           y = .93,
                           bjust = "center")

delay_group <- boxGrob(glue("Delay Groups"),
                       txt_gp = gpar(col = "black", fontsize = 16),
                       box_gp = gpar(fill = 0, col = 0),
                       x = .75,
                       y = .93,
                       bjust = "center")

group_am <- boxGrob(glue("AM-Control Group"),
                    txt_gp = gpar(col = "black", fontface = "bold"),
                    box_gp = gpar(fill = "#fee59d", col = "#fee59d"),
                    y = .8,
                    x = .15,
                    width = unit(0.175, "npc"),
                    height = unit(0.05, "npc"),
                    bjust = "center")

group_pm <- boxGrob(glue("PM-Control Group"),
                    txt_gp = gpar(col = "black", fontface = "bold"),
                    box_gp = gpar(fill = "#3276b4", col = "#3276b4"),
                    y = .8,
                    x = .35,
                    width = unit(0.175, "npc"),
                    height = unit(0.05, "npc"),
                    bjust = "center")

group_wake <- boxGrob(glue("Wake Group"),
                      txt_gp = gpar(col = "black", fontface = "bold"),
                      box_gp = gpar(fill = "#fff1ce", col = "#fff1ce"),
                      y = .8,
                      x = .65,
                      width = unit(0.175, "npc"),
                      height = unit(0.05, "npc"),
                      bjust = "center")

group_sleep <- boxGrob(glue("Sleep Group"),
                       txt_gp = gpar(col = "black", fontface = "bold"),
                       box_gp = gpar(fill = "#5e9cd3", col = "#5e9cd3"),
                       y = .8,
                       x = .85,
                       width = unit(0.175, "npc"),
                       height = unit(0.05, "npc"),
                       bjust = "center")

#start times

start_am <- boxGrob(glue("9AM Start"),
                    txt_gp = gpar(col = "black"),
                    box_gp = gpar(fill = 0, col = 0),
                    y = .7,
                    x = .15,
                    width = unit(0.1, "npc"),
                    height = unit(0.05, "npc"),
                    bjust = "top")

start_pm <- boxGrob(glue("9PM Start"),
                    txt_gp = gpar(col = "black"),
                    box_gp = gpar(fill = 0, col = 0),
                    y = .7,
                    x = .35,
                    width = unit(0.1, "npc"),
                    height = unit(0.05, "npc"),
                    bjust = "top")

start_wake <- boxGrob(glue("9AM Start"),
                      txt_gp = gpar(col = "black"),
                      box_gp = gpar(fill = 0, col = 0),
                      y = .7,
                      x = .65,
                      width = unit(0.1, "npc"),
                      height = unit(0.05, "npc"),
                      bjust = "top")

start_sleep <- boxGrob(glue("9PM Start"),
                       txt_gp = gpar(col = "black"),
                       box_gp = gpar(fill = 0, col =0),
                       y = .7,
                       x = .85,
                       width = unit(0.1, "npc"),
                       height = unit(0.05, "npc"),
                       bjust = "top")

#Test times

test_am <- boxGrob(glue("9AM Test"),
                   txt_gp = gpar(col = "black"),
                   box_gp = gpar(fill = 0, col = 0),
                   y = .320,
                   x = .15,
                   width = unit(0.1, "npc"),
                   height = unit(0.05, "npc"),
                   bjust = "top")

test_pm <- boxGrob(glue("9PM Test"),
                   txt_gp = gpar(col = "black"),
                   box_gp = gpar(fill = 0, col = 0),
                   y = .320,
                   x = .35,
                   width = unit(0.1, "npc"),
                   height = unit(0.05, "npc"),
                   bjust = "top")

test_wake <- boxGrob(glue("9PM Test"),
                     txt_gp = gpar(col = "black"),
                     box_gp = gpar(fill = 0, col = 0),
                     y = .320,
                     x = .65,
                     width = unit(0.1, "npc"),
                     height = unit(0.05, "npc"),
                     bjust = "top")
test_sleep <- boxGrob(glue("9AM Test"),
                      txt_gp = gpar(col = "black"),
                      box_gp = gpar(fill = 0, col =0),
                      y = .320,
                      x = .85,
                      width = unit(0.1, "npc"),
                      height = unit(0.05, "npc"),
                      bjust = "top")

study_phase <- boxGrob(glue("Study Phase"),
                       txt_gp = gpar(col = "black", fontface = "bold", fontsize = 20),
                       box_gp = gpar(fill = 0, col = 0),
                       x = .5,
                       y = .6,
                       bjust = "center")

language_check <- boxGrob(glue("Language/Attention Check"),
                          txt_gp = gpar(col = "black", fontsize = 13),
                          box_gp = gpar(fill = 0, col = 0),
                          x = .5,
                          y = .5,
                          bjust = "center")

stanford_scale <- boxGrob(glue("Stanford Sleepiness Scale (SSS)"),
                          txt_gp = gpar(col = "black", fontsize = 13),
                          box_gp = gpar(fill = 0, col = 0),
                          x = 0.15,
                          y = 0.5,
                          width = 0.3,
                          bjust = "center",
                          just = "right")

view_wordlist <- boxGrob(glue("View 20 DRM wordlists"),
                         txt_gp = gpar(col = "black", fontsize = 13),
                         box_gp = gpar(fill = 0, col = 0),
                         x = 0.85,
                         y = 0.5,
                         width = 0.3,
                         bjust = "center",
                         just = "left")

test_phase <- boxGrob(glue("Test Phase"),
                      txt_gp = gpar(col = "black", fontface = "bold", fontsize = 20),
                      box_gp = gpar(fill = 0, col = 0),
                      x = .5,
                      y = .22,
                      bjust = "center")

survey <- boxGrob(glue("Survey"),
                  txt_gp = gpar(col = "black", fontsize = 13),
                  box_gp = gpar(fill = 0, col = 0),
                  x = .5,
                  y = .12,
                  bjust = "center")

sss <- boxGrob(glue("SSS"),
               txt_gp = gpar(col = "black", fontsize = 13),
               box_gp = gpar(fill = 0, col = 0),
               x = 0.257,
               y = 0.12,
               width = 0.1,
               bjust = "center",
               just = "right")

free_recall <- boxGrob(glue("Free Recall"),
                       txt_gp = gpar(col = "black", fontsize = 13),
                       box_gp = gpar(fill = 0, col = 0),
                       x = 0.743,
                       y = 0.12,
                       width = 0.1,
                       bjust = "center",
                       just = "right")

#function for the braces
# x1...y2 are the ends of the brace as coordinates
# for upside down braces, x1 > x2 and y1 > y2
# h is the height -> how much higher the peak is than the body
curly_brace <- function(x1,y1,x2,y2,h) {
  x2 <- x2-x1; y2 <- y2-y1
  v1 <- viewport(x=x1,y=y1,width=sqrt(x2^2+y2^2),
                 height=h,angle=180*atan2(y2,x2)/pi,
                 just=c("left","bottom"),gp=gpar(col="lightblue", lwd = 2))
  pushViewport(v1)
  grid.curve(x2=0,y2=0,x1=.125,y1=.5,curvature=.5)
  grid.move.to(.125,.5)
  grid.line.to(.375,.5)
  grid.curve(x1=.375,y1=.5,x2=.5,y2=1,curvature=.5)
  grid.curve(x2=1,y2=0,x1=.875,y1=.5,curvature=-.5)
  grid.move.to(.875,.5)
  grid.line.to(.625,.5)
  grid.curve(x2=.625,y2=.5,x1=.5,y1=1,curvature=.5)
  popViewport()
}

#create chart

pdf(file = "Figures/expprocedure.pdf",
    width = 10,
    height = 5)

#add the images
image_sleep <- magick::image_read("Figures/sleep.tif")
image_wake <- magick::image_read("Figures/wake.tif")

ggdraw() +
  draw_image(image_sleep,
             x = 0.87,
             y = 0.4,
             width = 0.05,
             height = 0.05) +
  draw_image(image_wake,
             x = 0.67,
             y = 0.4,
             width = 0.05,
             height = 0.05)

#add the boxes
immediate_group
delay_group

group_am
group_pm
group_wake
group_sleep

start_am
start_pm
start_wake
start_sleep

test_am
test_pm
test_wake
test_sleep

#draw big box around centre part
grid.polygon(x = c(0.01,  0.99,  0.99,  0.01),
             y = c(0.65, 0.65, 0.45, 0.45),
             id = rep(1, 4),
             gp = gpar(col = "black", lwd = 4))

#draw smaller box around lower part
grid.polygon(x = c(0.15, 0.85, 0.85, 0.15),
             y = c(0.27, 0.27, 0.07, 0.07),
             id = rep(1, 4),
             gp = gpar(col = "black", lwd = 4))

#draw arrows

#from box AM-control
grid.polygon(x = c(0.130, 0.170, 0.170, 0.185, 0.150, 0.115, 0.130),
             y = c(0.775, 0.775, 0.725, 0.725, 0.700, 0.725, 0.725),
             id = rep(1, 7),
             gp = gpar(fill = "#fee59d", col = "black"), draw = TRUE)

#from box PM-control
grid.polygon(x = c(0.330, 0.370, 0.370, 0.385, 0.350, 0.315, 0.330),
             y = c(0.775, 0.775, 0.725, 0.725, 0.700, 0.725, 0.725),
             id = rep(1, 7),
             gp = gpar(fill = "#3276b4", col = "black"), draw = TRUE)

#from box Wake Group
grid.polygon(x = c(0.630, 0.670, 0.670, 0.685, 0.650, 0.615, 0.630),
             y = c(0.775, 0.775, 0.725, 0.725, 0.700, 0.725, 0.725),
             id = rep(1, 7),
             gp = gpar(fill = "#fff1ce", col = "black"), draw = TRUE)

#from box Sleep Group
grid.polygon(x = c(0.830, 0.870, 0.870, 0.885, 0.850, 0.815, 0.830),
             y = c(0.775, 0.775, 0.725, 0.725, 0.700, 0.725, 0.725),
             id = rep(1, 7),
             gp = gpar(fill = "#5e9cd3", col = "black"), draw = TRUE)

#lower row of arrows

#from big box to 9am left
grid.polygon(x = c(0.130, 0.170, 0.170, 0.185, 0.150, 0.115, 0.130),
             y = c(0.447, 0.447, 0.350, 0.350, 0.320, 0.350, 0.350),
             id = rep(1, 7),
             gp = gpar(fill = "#fee59d", col = "black"), draw = TRUE)

#from big box to 9pm left
grid.polygon(x = c(0.330, 0.370, 0.370, 0.385, 0.350, 0.315, 0.330),
             y = c(0.447, 0.447, 0.350, 0.350, 0.320, 0.350, 0.350),
             id = rep(1, 7),
             gp = gpar(fill = "#3276b4", col = "black"), draw = TRUE)

#from big box to 9am right
grid.polygon(x = c(0.630, 0.670, 0.670, 0.685, 0.650, 0.615, 0.630),
             y = c(0.447, 0.447, 0.350, 0.350, 0.320, 0.350, 0.350),
             id = rep(1, 7),
             gp = gpar(fill = "#fff1ce", col = "black"), draw = TRUE)

#from big box to 9pm right
grid.polygon(x = c(0.830, 0.870, 0.870, 0.885, 0.850, 0.815, 0.830),
             y = c(0.447, 0.447, 0.350, 0.350, 0.320, 0.350, 0.350),
             id = rep(1, 7),
             gp = gpar(fill = "#5e9cd3", col = "black"), draw = TRUE)

study_phase
language_check
stanford_scale
view_wordlist

test_phase
survey
sss
free_recall

#arrows in big box

#from Stanford scale to check
grid.polygon(x = c(0.300, 0.355, 0.355, 0.375, 0.355, 0.355, 0.300),
             y = c(0.510, 0.510, 0.530, 0.500, 0.470, 0.490, 0.490))

#from check to DRM
grid.polygon(x = c(0.625, 0.680, 0.680, 0.700, 0.680, 0.680, 0.625),
             y = c(0.510, 0.510, 0.530, 0.500, 0.470, 0.490, 0.490))

#arrows in small box

#from SSS to Survey
grid.polygon(x = c(0.365, 0.420, 0.420, 0.440, 0.420, 0.420, 0.365),
             y = c(0.130, 0.130, 0.150, 0.120, 0.090, 0.110, 0.110))

#from Survey to free recall
grid.polygon(x = c(0.560, 0.615, 0.615, 0.635, 0.615, 0.615, 0.560),
             y = c(0.130, 0.130, 0.150, 0.120, 0.090, 0.110, 0.110))

#curly braces

curly_brace(0.05, 0.84, 0.45, 0.84, 0.05)
curly_brace(0.55, 0.84, 0.95, 0.84, 0.05)

#delays
grid.text("No\nDelay",
          x = .15,
          y = .39,
          just = c("center"),
          rot = 90,
          gp = gpar(fontsize = 10))

grid.text("No\nDelay",
          x = .35,
          y = .39,
          just = c("center"),
          rot = 90,
          gp = gpar(fontsize = 10))

grid.text("12-hr\nDelay",
          x = .65,
          y = .39,
          just = c("center"),
          rot = 90,
          gp = gpar(fontsize = 10))

grid.text("12-hr\nDelay",
          x = .85,
          y = .39,
          just = c("center"),
          rot = 90,
          gp = gpar(fontsize = 10))

#underline study phase
grid.lines(x = c(0.418,0.582),
           y = c(0.57),
           gp = gpar(lwd = 3))

#underline test phase
grid.lines(x = c(0.428, 0.572),
           y = c(0.19),
           gp = gpar(lwd = 3))

dev.off()

### 2023-07-25 manuscript.tex
Test autogenerate of the procedure figure

Here, we asked whether recall probability of a lure (e.g., doctor) is predicted by the number of corresponding list items being recalled (e.g., nurse, hospital, sick), and if it does, whether it differs between the Sleep and Wake groups. These questions help shed light on the degree to which sleep increases lure recall via processes such as retrieval-induced generalisation\footnote{Retrieval-induced generalisation: Retrieval of one word cueing retrieval of a related word (e.g., Berens \& Bird, 2017)} or gist abstraction\footnote{Gist abstraction: Extraction of the central or essential meaning of learned information}, as these processes may predict a different degree of interdependence between lure and veridical recall. If sleep (vs.~wake) promoted retrieval-induced generalisation, lure and veridical recall should become more strongly correlated with each other after sleep, because better veridical recall for a set of studied words may generalise to the corresponding critical lure (or vice versa). On the other hand, if sleep (vs.~wake) promoted gist abstraction, lure recall may become less related to memories for the corresponding list items, because theories of gist abstraction such as iOtA may predict that sleep would selectively boost the overlapping gist memory (i.e., the lure) but not necessarily the specific studied words.

@@

As per the DRM literature, responses that were neither the studied list words nor the critical lures were classified as (non-critical) intrusions. For instance, our participants studied nurse, sick, lawyer, medicine, health, hospital, dentist, and physician, with doctor as the critical lure. Responses such as clinic and coconut would both be considered intrusions, but clearly, clinic is semantically more related to the studied list words than coconut. In other words, there is much diversity within the intrusion data. Here, we explored whether our four groups differed in terms of the semantic distance between intrusions and critical lures, as indexed by pre-trained semantic spaces (ukWaC; {[}\protect\hyperlink{ref-baroni2009a}{45}{]}) derived from word2vec {[}\protect\hyperlink{ref-guenther2015a}{46}{]}. We reasoned that since lure recall was greater in our Sleep (vs.~Wake) participants, the intrusions produced by these participants could be more related to the lures in semantic space (e.g., {[}\protect\hyperlink{ref-mak2023a}{15}{]}). To test this, we computed the cosine similarities between each intrusion and each of the 20 critical lures (see Table \ref{tab:table8} for an illustration). The intrusion-lure pair with the highest cosine similarity (i.e., the nearest neighbour) was used for this analysis.

>>>>

Here, we asked whether recall probability of a lure (e.g., \emph{doctor}) is predicted by the number of corresponding list items being recalled (e.g., \emph{nurse}, \emph{hospital}, \emph{sick}), and if it does, whether it differs between the Sleep and Wake groups. These questions help shed light on the degree to which sleep increases lure recall via processes such as retrieval-induced generalisation\footnote{Retrieval-induced generalisation: Retrieval of one word cueing retrieval of a related word (e.g., Berens \& Bird, 2017)} or gist abstraction\footnote{Gist abstraction: Extraction of the central or essential meaning of learned information}, as these processes may predict a different degree of interdependence between lure and veridical recall. If sleep (vs.~wake) promoted retrieval-induced generalisation, lure and veridical recall should become more strongly correlated with each other after sleep, because better veridical recall for a set of studied words may generalise to the corresponding critical lure (or vice versa). On the other hand, if sleep (vs.~wake) promoted gist abstraction, lure recall may become less related to memories for the corresponding list items, because theories of gist abstraction such as iOtA may predict that sleep would selectively boost the overlapping gist memory (i.e., the lure) but not necessarily the specific studied words.

@@

As per the DRM literature, responses that were neither the studied list words nor the critical lures were classified as (non-critical) intrusions. For instance, our participants studied \emph{nurse}, \emph{sick}, \emph{lawyer}, \emph{medicine}, \emph{health}, \emph{hospital}, \emph{dentist}, and \emph{physician}, with \emph{doctor} as the critical lure. Responses such as \emph{clinic} and \emph{coconut} would both be considered intrusions, but clearly, \emph{clinic} is semantically more related to the studied list words than \emph{coconut}. In other words, there is much diversity within the intrusion data. Here, we explored whether our four groups differed in terms of the semantic distance between intrusions and critical lures, as indexed by pre-trained semantic spaces (ukWaC; {[}\protect\hyperlink{ref-baroni2009a}{45}{]}) derived from word2vec {[}\protect\hyperlink{ref-guenther2015a}{46}{]}. We reasoned that since lure recall was greater in our Sleep (vs.~Wake) participants, the intrusions produced by these participants could be more related to the lures in semantic space (e.g., {[}\protect\hyperlink{ref-mak2023a}{15}{]}). To test this, we computed the cosine similarities between each intrusion and each of the 20 critical lures (see Table \ref{tab:table8} for an illustration). The intrusion-lure pair with the highest cosine similarity (i.e., the nearest neighbour) was used for this analysis.
